---
title: "Assignment 1"
author: "Folorunsho Atanda"
date: "2023-09-04"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objective: The objective of this assignment is to take data set and make a subset
of it. For my task I took the data set that shows the approval rating of
Trump and Biden based on how they handle the covid pandemic

```{r import libraries}
## import data.table library
library(data.table)
## import ggplot library
library(ggplot2)
```

```{r data}
## import covid approval poll csv 
url <- "https://raw.githubusercontent.com/folushoa/Data-Science/Data-607/covid_approval_polls.csv"
covid_approval_poll <- fread(input = url, sep = ",", header = TRUE)
```

```{r subset}
## subset of covid approval status focusing on Trump
trump_approval_poll <- subset(covid_approval_poll, 
                              subject == "Trump" & party == "all", 
                              select = c(start_date, end_date, approve, 
                                         disapprove, sample_size))

## subset of covid approval status focusing on Biden
biden_approval_poll <- subset(covid_approval_poll, 
                              subject == "Biden" & party == "all", 
                              select = c(start_date, end_date, approve, 
                                         disapprove, sample_size))
```

```{r Trump plot}
## plot of Trump approval and disapproval
trump_poll_plot <- ggplot(data = trump_approval_poll, aes(x = end_date, 
                   y = approve)) + geom_line(color = 'blue') + labs(x = "Month",
                   y = "Ratio", title = "Trump Approval Rating")
trump_poll_plot <- trump_poll_plot + geom_line(aes(x = end_date, 
                   y = disapprove), color = 'red')
trump_poll_plot <- trump_poll_plot + 
  scale_color_manual(name = "Approval Status", breaks = c('Approve', 'Disapprove'),
  values = c('Approve' =  'blue', 'Disapprove' = 'red'))
trump_poll_plot
```

Observation: Examining the preceding graph illustrating Trump's approval ratings 
(indicated by the color blue) and disapproval ratings 
(indicated by the color red), it becomes evident that at the onset of the 
pandemic, Trump enjoyed a favorable approval rating. However, this dynamic 
shifted in April, as we observe a notable rise in his disapproval ratings.

When President Biden assumed office in January 2021, Trump's disapproval rating 
surpassed his approval rating.

```{r Biden Plot}
## plot of Biden approval and disapproval
biden_poll_plot <- ggplot(data = biden_approval_poll, aes(x = end_date, 
                   y = approve)) + geom_line(color = 'blue') + labs(x = "Month",
                   y = "Ratio", title = "Biden Approval Rating")
biden_poll_plot <- biden_poll_plot + geom_line(aes(x = end_date, 
                   y = disapprove), color = 'red')
biden_poll_plot <- biden_poll_plot + 
  scale_color_manual(name = "Approval Status", breaks = c('Approve', 'Disapprove'),
                     values = c('Approve' =  'blue', 'Disapprove' = 'red'))
biden_poll_plot
```

Observation: Analyzing the aforementioned chart depicting Biden's approval 
ratings (represented in blue) and disapproval ratings (represented in red), it 
becomes evident that at the commencement of the pandemic, the majority of 
individuals endorsed Biden's pandemic management. Nevertheless, as of October 
2022, the balance between approval and disapproval became roughly equal. This 
equilibrium persisted until the conclusion of the polling period in 
September 2022.