---
title: "Assignment2"
author: "Folorunsho Atanda"
date: "2023-09-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required library
```{r}
library(DBI)
library(RMySQL)
library(dplyr)
library(dotenv)
```

## Create database connection
```{r}
dotenv::load_dot_env(file = "sql_pw.env")
my_sql_pw <- Sys.getenv("MYSQL_PW")

db <- dbConnect(MySQL(),
      user = "root",
      password = my_sql_pw,
      dbname = "assignment_2",
      host = "localhost",
      port = 3306)

popular_movie_rating <- dbGetQuery(db, 
                        "select * from assignment_2.popular_movie_rating")

PMR_df <- as.data.frame((popular_movie_rating))
PMR_df 
```

## Table summary
```{r}
summary(PMR_df)
```

## Table glimpse
```{r}
glimpse(PMR_df)
```

## Create filtered tables based on movies.
## First factor the movie name column to see the unique categories
```{r}
factor(PMR_df$Movie_Name) %>%levels()
```

## Now filter 
```{r}
barbie_PMR_df <- PMR_df %>% filter(Movie_Name == "Barbie" & 
                            is.na(Movie_Rating) == FALSE)
barbie_PMR_df 

indiana_jones_PMR_df <- PMR_df %>% filter(Movie_Name == "Indiana Jones" & 
                                   is.na(Movie_Rating) == FALSE)
indiana_jones_PMR_df 

oppenheimer_PMR_df <- PMR_df %>% filter(Movie_Name == "Oppenheimer" & 
                                 is.na(Movie_Rating) == FALSE)
oppenheimer_PMR_df 

tmnt_PMR_df <- PMR_df %>% filter(Movie_Name == "TMNT: Mutant Mayhem" & 
                          is.na(Movie_Rating) == FALSE)
tmnt_PMR_df 

nimona_PMR_df <- PMR_df %>% filter(Movie_Name == "Nimona" & 
                            is.na(Movie_Rating) == FALSE)
nimona_PMR_df 

meg2_PMR_df <- PMR_df %>% filter(Movie_Name == "Meg 2" & 
                          is.na(Movie_Rating) == FALSE)
meg2_PMR_df 
```

## To find the central value of the ratings we shall use median.
## My reason for using median is because the number of reviews for each
## movie is different now that we have removed **NA** values.
```{r}
barbie_PMR_df_median <- barbie_PMR_df %>% summarise(median(Movie_Rating))
ij_PMR_df_median <- indiana_jones_PMR_df %>% summarise(median(Movie_Rating))
opp_PMR_df_median <- oppenheimer_PMR_df %>% summarise(median(Movie_Rating))
tmnt_PMR_df_median <- tmnt_PMR_df %>% summarise(median(Movie_Rating))
nimona_PMR_df_median <- nimona_PMR_df %>% summarise(median(Movie_Rating))
meg2_PMR_df_median <- meg2_PMR_df %>% summarise(median(Movie_Rating))
```

## Place median (overall rating) in a table
```{r}
popular_movie_overall_rating <- data.frame(
  Movie_Name = c("Barbie", "Indiana Jones", "Meg 2", "Nimona", "Oppenheimer",
                 "TMNT: Mutant Mayhem"), 
  Overall_Rating = c(barbie_PMR_df_median$`median(Movie_Rating)`, 
                     ij_PMR_df_median$`median(Movie_Rating)`, 
                     meg2_PMR_df_median$`median(Movie_Rating)`,
                     nimona_PMR_df_median$`median(Movie_Rating)`, 
                     opp_PMR_df_median$`median(Movie_Rating)`, 
                     tmnt_PMR_df_median$`median(Movie_Rating)`)
)
popular_movie_overall_rating 
```

