---
title: "Data605 HW2"
author: "Folorunsho Atanda"
date: "2024-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressMessages(
  library(tidyverse)
)
```

# Problem Set 1.1
## Show that **_A^T^A $\neq$ AA^T^_** in general form
Let _A_ be a _m by n_ matrix such that _m $\neq$ n. Therefore _A^T^A will equal
a matrix with dimension _n by n_. While _AA^T^_ will equal a matrix with 
dimension _m by m_. Since _m $\neq$ n_, then _A^T^A $\neq$ AA^T^_. QED 

For example:
```{r 1.1 Demonstration}
A <- matrix(seq(1, 6, by = 1), nrow = 3, byrow = TRUE)
# Transpose of A left multiply A
t(A) %*% A
# A left multiply Transpose of A
A %*% t(A)
```

# Problem Set 1.2
## For a special type of square matrix **_A_**, we get **_A^T^A $=$ AA^T^_**. 
## Under what conditin is this true?
For a square matrix _A_, _A^T^A $=$ AA^T^_ is true if _A_ is a **symmetrical**
**matrix**, i.e. A $=$ A^T^

For example:
```{r 1.2 Demonstration}
content <- c(3, 4, 4, 2)
A <- matrix(content, nrow = 2, byrow = TRUE)
# Compare A with it Transpose
A == t(A) # Both matrix are equal

# Compare Transpose of A left multiply A with # A left multiply Transpose of A
t(A) %*% A == A %*% t(A) # Both matrix are equal
```
Therefore if A $=$ A^T^ then AA^T^ $=$ A^T^A


# Problem Set 2
```{r LU Factorization}
lu_factorization <- function(A) {
  # Check if the matrix is square
  if (ncol(A) != nrow(A)) {
    return("Input matrix must be square for LU factorization.")
  }
  
  n <- nrow(A)
  
  # Initialize L and U matrices
  L <- diag(n)
  U <- matrix(0, n, n)
  
  for (k in 1:n) {
    # Populate L and U matrices
    U[k, k] <- A[k, k]
    
    if(k + 1 <= n){
      for (i in (k + 1):n){
        L[i, k] <- A[i, k]/U[k, k]
        U[k, i] <- A[k, i]
      }
      
      # Update rest of matrix accordingly
      for (i in (k + 1):n){
        for (j in (k + 1):n){
          A[i, j] <- A[i, j] - (L[i, k] %*% U[k, j])
        }
      }
    }
  }
  # Return the L and U
  return(list(L = L, U = U))
}

# Example usage
A <- matrix(c(2, 3, 1, 5, 6, 13, 5, 19, 2, 19, 10, 23, 4, 10, 11, 31), 
            nrow = 4, 
            byrow = TRUE)

# Perform LU factorization
lu_result <- lu_factorization(A)

# Display the result
print("Original Matrix A:")
print(A)

print("Matrix L:")
print(lu_result$L)

print("Matrix U:")
print(lu_result$U)

# Compare LU factorization with original A matrix
(lu_result$L %*% lu_result$U) == A
```

