---
title: "Assignment_3"
author: "Folorunsho Atanda"
date: "2023-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
```

1.  Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset,
provide code that identifies the majors that contain either "DATA" or
"STATISTICS"

```{r}
url = "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
major_list = fread(url, header = TRUE)
data_statistic_major <- str_view(major_list$Major, "DATA|STATISTICS")
data_statistic_major
```

2. Write code that transforms the data below

```{r}
fg <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'

pattern <- '\\"[a-zA-Z]+\\ ?[a-zA-Z]+?\\"'
fruit_and_vegetable <- unlist(str_extract_all(fg, pattern))
fruits <- str_remove_all(fruit_and_vegetable, "\"")
fruits
```

3. Describe, in words, what these expressions will match:
a. **(.)\1\1**: This regular expression will match any string that has three
repeated characters, e.g.

```{r}
str_view(c("caaarve", "trinnne", "craske"), "(.)\\1\\1")
str_detect(c("caaarve", "trinnne", "craske"), "(.)\\1\\1")
```

b. **"(.)(.)\\2\\1"**: This regular expression will match any string that has
at least four character where from the first grouping the third character 
matches the second character and the fourth character matches the first. e.g.

```{r}
str_view(c("caa", "raaretar", "trrt", "abba", "bassat"), "(.)(.)\\2\\1")
str_detect(c("caa", "raaretar", "trrt", "abba", "bassat"), "(.)(.)\\2\\1")
```

c. **(..)\1**: This regular expression will match any string that has at least 
four character where from the first grouping the next two consecutive characters
match the first two preceding characters. e.g.

```{r}
str_view(c("caac", "raarar", "trtr", "abba", "sat"), "(..)\\1")
str_detect(c("caac", "raarar", "trtr", "abba", "sat"), "(..)\\1")
```

d. **"(.).\\1.\\1"**: This regular expression will match any string that has at 
least five characters where from the first grouping the third and fifth 
characters match the first characters. The second and fourth character don't 
need to match. e.g.

```{r}
str_view(c("cacrcab", "rbti", "grgrg", "btvqvrv"), "(.).\\1.\\1")
str_detect(c("cacrcab", "rbti", "grgrg", "btvqvrv"), "(.).\\1.\\1")
```

e. **"(.)(.)(.).*\\3\\2\\1"**: This regular expression will match any strings 
that has at least six characters where from the first three grouping there might
be a bunch of characters but then the first three grouping repeat themselves but
in reverse order. e.g.

```{r}
str_view(c("abccba", "gretcdterab", "gvbe", "qrstuv"), "(.)(.)(.).*\\3\\2\\1")
str_detect(c("abccba", "gretcdterab", "gvbe", "qrstuv"), "(.)(.)(.).*\\3\\2\\1")
```

4. Construct regular expressions to match words that:
a. Start and end with the same character: **^([a-zA-Z]).*\1$** e.g.

```{r}
str_view(c("apple", "anana", "abba", "creerp"), "^([a-zA-Z]).*\\1$")
```

b. Contain a repeated pair of letters: **[a-zA-Z][a-zA-Z]).*\1** e.g.

```{r}
str_view(fruits, "([a-zA-Z][a-zA-Z]).*\\1")
```

c. Contain one letter repeated in at least three places: 
**([a-zA-Z]).*\1.*\1.** e.g.

```{r}
str_view(fruits, "([a-zA-Z]).*\\1.*\\1.")
```

