---
title: "Assignment_5"
author: "Folorunsho Atanda"
date: "2023-10-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(tidyverse)
library(data.table)
```

# Read created cvs from github

```{r}
url <- "https://raw.githubusercontent.com/folushoa/Data-Science/Data-607/Assignment%205/Assignment%205.csv"
airline_data <- fread(url, na.strings = "")
airline_data
```

# Tidy data

```{r}
airline_data <- airline_data %>% 
  fill(Airline, .direction = "down")
airline_data

airline_data <- airline_data %>%
  pivot_longer(
    cols = !c(Airline, Status),
    names_to = "City",
    values_to = "Count"
  )
airline_data
```

# Analyze data

Compare arrival delays of the two airlines

```{r}
airline_data_delay <- airline_data %>% 
  filter(Status == "Delayed") %>% 
  group_by(Airline) %>% 
  summarise(
    "Total Count" = sum(Count),
  ) %>% 
  mutate(
    "Proportion" = signif(`Total Count`/sum(`Total Count`), digits = 4)
  )
airline_data_delay

airline_data_delay %>%
  ggplot(aes(x = `Airline`, y = `Proportion`)) +
  geom_bar(stat = "identity") 
```

**Conclusion**:
From the table and the plot we see that Alaska airline had less delays than 
AM West, 0.389 to 0.611

Compare arrival delays by city

```{r}
airline_data_delay <- airline_data %>% 
  filter(Status == "Delayed") %>% 
  group_by(Airline) %>% 
  mutate(
    "Proportion" = signif(`Count`/sum(`Count`), digits = 3)
  )
airline_data_delay

airline_data_delay %>%
  ggplot(aes(x = `Airline`, y = `Proportion`)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~City)
```

**Conclusion**: 
From the plots by City, AM West arrives late more times than Alaska airline. So 
I would conclude that it is better to travel by AM West than Alaska.