---
title: "Assignment 7"
author: "Folorunsho Atanda"
date: "2023-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load library
```{r}
library(data.table)
library(tidyverse)
library(rvest) #for reading HTML file into R
library(jsonlite) #for reading JSON file into R
library(xml2)
```

# HTML Format
Lets read a HTML table into R

## Save HTML URL into R
```{r}
url <- "https://raw.githubusercontent.com/folushoa/Data-Science/Data-607/Assignment%207/books.html"
html <- read_html(url)
```

## Convert to data frame
```{r}
html_table <- html %>% 
  html_element("table") %>% 
  html_table()

html_table
```

# JSON Format
Lets read a JSON table into R

## Save JSON url into R
```{r}
url <- "https://raw.githubusercontent.com/folushoa/Data-Science/Data-607/Assignment%207/books.json"
json_table <- fromJSON(url, flatten = TRUE) #reads JSON table and converts to
                                            #data frame

json_table
```

# XML Format
Lets read a XML table into R

## Save XML URL into R
```{r}
url <- "https://raw.githubusercontent.com/folushoa/Data-Science/Data-607/Assignment%207/books.xml"
xml <- read_xml(url)
```

## Convert to data frame
```{r}
name_nodes <- xml %>% #extract all name nodes
  xml_find_all(".//book") %>% 
  xml_find_all(".//Name")
  
genre_nodes <- xml %>% #extract all genre nodes
  xml_find_all(".//book") %>% 
  xml_find_all(".//Genre")

author_nodes <- xml %>% #extract all author nodes
  xml_find_all(".//book") %>% 
  xml_find_all(".//Author")

# Extract text from the nodes
name <- xml_text(name_nodes)
genre <- xml_text(genre_nodes)
author <- xml_text(author_nodes)

# Put into a data frame
xml_table <- tibble(Name = name, Genre = genre, Author = author)

xml_table
```

The HTML and SML to data frame, data frames are the same, tibble. However, the 
JSON to data frame is a data frame. All in all they are all data frames.